\Large \bf DRAFT: Direct-style process creation on Linux

### 2020-05-23 research/paper/paper.tex
paper

As a result, though implementations in native code and in the kernel would likely remove most of this overhead,
we have chosen to not invest effort into optimizing process creation at this micro-level,
to preserve implementation simplicity;
such optimization is reserved for future work.
\subsection{Usage in the real world}\label{realworld}
At \twosigma we have used \texttt{rsyscall} and direct-style process creation
as a component of an internal distributed system deployment library, written in Python.
This library is in use in production,
and is extensively used as part of testing and development.

Each service deployed by this library is started with custom code,
written using direct-style process creation.
This allows us to pass down file descriptors and initialize process attributes
according to each program's needs;
which in turn has let us make optimizations which previously required too much program-specific customization.

For example,
we are able to bootstrap connections over our internal shared memory transports over passed-down file descriptors,
and our management interfaces listen on passed-down sockets.
This allows substantial parallelism in the startup of services.\cite{socketactivation}
Taking one representative example,
one test built with an old library took an average of 221 seconds to run,
mostly spent in system startup time as processes started up serially.
With parallelized system startup, the same test takes an average of 16 seconds.

We've found that programmers
are able to use file descriptor passing as a normal feature of their development process with relative ease.
Programmers without substantial prior experience with Unix programming
are able to write code to support new services
nd pass down file descriptors to the program without encountering significant issues.

We also flexibly use namespaces where appropriate.
For example, for legacy applications which directly start subprocesses,
we prevent process leaking with a pid namespace.

>>>>

As a result, though implementations in native code and in the kernel would likely remove most of this overhead,
we have chosen to not invest effort into optimizing process creation at this micro-level,
to preserve implementation simplicity;
such optimization is reserved for future work.

\subsection{Microbenchmarks for nested processes}\label{clone_bench}
\begin{figure}[h!]
\centering
 \includegraphics[width=0.5\textwidth]{clone_bench}
 \caption{Time to create different processes, by variety of parent process}
 \label{fig:clone_bench}
\end{figure}

As demonstrated in sections \ref{pidns} and \ref{fuse},
\texttt{rsyscall} supports nested operation.
That is, we can spawn a child from one of our children;
this can impact performance.
\texttt{clone} also supports specifying a variety of flags to create new namespaces when creating the child,
which can also impact performance.

We've benchmarked process creation using the same benchmark as in section \ref{subprocess_bench}
from a variety of parent processes: the local process,
a child (created with \texttt{clone()}),
a grandchild (created with \texttt{clone().clone()}),
or a child in a different namespace (created with \verb=clone(CLONE_NEWNS|CLONE_NEWPID)=).
We've also specified different combinations of flags for the process being created in the benchmark.
The results are in figure \ref{fig:clone_bench}.
We don't include the Python standard library in these benchmarks
as there's no way to specify \texttt{CLONE} flags in the Python stdlib.

Process creation in a child has a significant performance impact,
primarily because of the overhead of performing syscalls in a child.
Performance of process creation in a grandchild is the same as in a child,
because there is no additional overhead for syscalls in a grandchild relative to a child.
Benchmarking \texttt{getpid} in a grandchild
confirms that the average time per \texttt{getpid} call is still 561 microseconds on our benchmark setup.

The use of namespaces also has a significant impact,
both when the parent process is in a namespace, and when the child process is created in a namespace.
We perform the benchmark with no clone flags, and with the common clone flags
\verb|CLONE_NEWNS| and \verb|CLONE_NEWPID|, first on their own and then together.
This benchmark essentially measures the performance of the Linux namespaces system, not the work in this paper,
but it is illustrative to compare it to the performance of \texttt{rsyscall}.
Note that the cost of creating a new child process from a parent process that is already inside a namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} parent processes)
is around 2 milliseconds, comparable to the cost of running with \texttt{rsyscall}.
Note also that the overhead of creating a new process inside a new namespace
(the difference between the \texttt{clone()} and \texttt{clone(NS|PID)} child processes)
is around 15 milliseconds; subtantially greater than the overhead of \texttt{rsyscall}.
\subsection{Usage in the real world}\label{realworld}
At \twosigma we have used \texttt{rsyscall} and direct-style process creation
to implement an internal distributed system deployment library, written in Python,
which we'll refer to here as Toplevel.
Toplevel is in use in production,
and is extensively used as part of testing and development.
Toplevel consists of a collection of functions and modules
for each component in our system,
which start up components using direct-style process creation,
all from a single Python parent process.

One of the major benefits of direct-style process creation has been easy use of file descriptor inheritance.
Our experience has been that with fork-style or spawn-style interfaces,
all file descriptor logic must be centralized,
whereas file descriptors in a new process can be built up over time using direct-style,
simplifying the implementation.
We've found that even programmers without substantial prior experience with Unix programming
are able to use file descriptor passing as a normal feature of their development process with relative ease,
writing code to support new programs and passing down file descriptors to those programs without issue.

Easy file descriptor passing has, in turn, allowed us to heavily use socket activation techniques
to start up services in parallel,
substantially speeding up system startup.\cite{socketactivation}
We are able to bootstrap connections over our internal shared memory transports
over pre-created, passed-down file descriptors,
and our management interfaces listen on passed-down sockets;
this allows starting up services in parallel, rather than in dependency order.
Benchmarking one test built with an old library and its equivalent built with Toplevel,
the old test took an average of 221 seconds to run,
mostly spent in system startup,
while the new test takes an average of 16 seconds.

Easy file descriptor inheritance has also allowed us to treat file descriptors as a uniform interface in Toplevel.
We pass around file descriptors freely inside Toplevel.
A given component can implemented with a separate process which has the file descriptor passed down,
running a program written in native code for performance,
or implemented in-process with a Python implementation for flexibility;
either is possible behind a uniform file descriptor interface.

We've also been able to create complicated pipelines in Toplevel,
as in section \ref{pipe}.

We've also made use of namespaces,
chiefly user namespaces and pid namespaces.
For legacy applications which directly start subprocesses,
we prevent process leaking with a pid namespace.
This is part of strategy of noninterference and cleanup
that ensures that multiple users of Toplevel
can coexist simultaneously on the same host,
even if running the same system.

Toplevel also makes uses of other features of \texttt{rsyscall} which are not covered in this paper.
These include the ability to operate on remote hosts through capabilities for remote processes,
and the ability to perform system calls in parallel across a pool of processes.
These have significant synergy with direct-style process creation
in forming an overall powerful system.
